\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shapes interactions with floral visitors

### 2022-07-19 Manuscript/Lanuza_et_al_2022.tex
style edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  a4paper,
]{article}
\title{\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors}
\author{}
\date{\vspace{-2.5em}}

\usepackage{amsmath,amssymb}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={ Covariation among reproductive traits in flowering plants determine interactions with floral visitors},
  colorlinks=true,
  linkcolor={RoyalBlue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={RoyalBlue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\pagestyle{plain}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno} % add 
\linenumbers % turns line numbering on

@@

Plants have an enormous variation of life history strategies and trait combinations. However, there is strong evidence that there are evolutionary and physiological constraints that limit the number of plant ecological strategies. Despite recent advances in the understanding of plant trait correlations, reproductive traits are often not considered despite its key role in shaping both plant life-history strategies and plants interactions with floral visitors. Here, we investigate the reproductive spectrum of flowering plants and how this drive the establishment of interactions with their floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over 50\% of all trait variation was explained by the first two reproductive axes. Specifically, the first axis indicated the presence of a negative correlation between flower number and flower size while the second axis showed a positive correlation between pollinator dependence and floral display. These different reproductive axes determined important differences in the interaction level among floral visitor guilds but were insufficient to fully capture plant-pollinator associations. Our study shows the main reproductive trait correlations of flowering plants and highlights their relevance to understand plant-pollinator interactions in a global context.

@@

Flowering plants have an astonishing diversity of floral structures {[}\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}{]} that shape their interactions with animal pollinators {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}{]}. The variety of reproductive strategies followed by different plant species, from specialized reliance on a single pollinator, to full self-pollination, is incredibly rich. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints {[}\protect\hyperlink{ref-stearns1989}{5},\protect\hyperlink{ref-agrawal2020}{6}{]}. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits {[}\protect\hyperlink{ref-salguero2016}{7}--\protect\hyperlink{ref-song2022}{11}{]}, their study is often limited to a few number of reproductive traits that are rarely studied in combination. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other root {[}\protect\hyperlink{ref-laughlin2021}{12}{]}, leaf {[}\protect\hyperlink{ref-wright2004}{13}{]} or wood {[}\protect\hyperlink{ref-chave2009}{14}{]} related traits, there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the plant reproductive spectrum can help to further understand the different plant ecological strategies {[}\protect\hyperlink{ref-agrawal2020}{6}{]} and improve our knowledge on plant-pollinator associations {[}\protect\hyperlink{ref-roddy2021}{10}{]}.

With the recent availability of large trait databases {[}e.g., TRY \protect\hyperlink{ref-kattge2011}{15},~COMPADRE \protect\hyperlink{ref-salguero2015}{16}{]}, plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-diaz2016}{17}--\protect\hyperlink{ref-carmona2021}{19}{]}. However, most studies with a multitrait perspective have overlooked plant reproductive traits {[}\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-evojtko2020}{20}{]} and focused on trait correlations from other plant organs. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number {[}\protect\hyperlink{ref-sargent2007}{21},\protect\hyperlink{ref-kettle2011}{22}{]}, the positive association between flower size and outcrossing rate {[}\protect\hyperlink{ref-goodwillie2010}{23}{]} or the positive association between outcrossing rate and lifespan {[}\protect\hyperlink{ref-barrett2003}{24},\protect\hyperlink{ref-moeller2017}{25}{]}. Although these different trait correlations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships {[}\protect\hyperlink{ref-roddy2021}{10}{]}, we still lack empirical evidence that investigates jointly how these different reproductive trait associations shape plant life history strategies.

Despite this increasing number of macroecological studies that investigate specific plant reproductive traits {[}\protect\hyperlink{ref-moeller2017}{25}--\protect\hyperlink{ref-grossenbacher2017}{28}{]}, we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-sargent2008}{29}--\protect\hyperlink{ref-ruger2018}{31}{]}. The use of trait-based approaches {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{32}{]} and trait-matching analyses {[}\protect\hyperlink{ref-stang2009}{33},\protect\hyperlink{ref-bartomeus2016}{34}{]} has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles {[}e.g., specialists vs generalists \protect\hyperlink{ref-tur2013}{35}{]} and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators {[}\protect\hyperlink{ref-junker2013}{36},\protect\hyperlink{ref-coux2016}{37}{]}. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure {[}\protect\hyperlink{ref-stang2009}{33},\protect\hyperlink{ref-ibanez2012}{38}{]}. However, is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations {[}\protect\hyperlink{ref-tur2013}{35},\protect\hyperlink{ref-devaux2014}{39}{]}. Indeed, the pollination system of a great number of plant species remains unexplored {[}\protect\hyperlink{ref-dellinger2020}{4}{]}. Because both the species' morphology and mating system can determine the species' functional role in the pollination network and the combination of several traits has shown to increase the predictive power of the structure of ecological networks {[}\protect\hyperlink{ref-eklof2013}{40}{]}, an interesting novel approach is to investigate how a complete set of reproductive traits in the multidimensional trait space determine species interaction patterns {[}see \protect\hyperlink{ref-dehling2016}{41} for a morphological approach{]}. Thus, by exploring the plant reproductive spectrum of trait variation it is possible to delimit the different reproductive strategies and explore how these are associated with the different floral visitors.

Here, we aim to progress knowledge on the reproductive trait covariation patterns and their association with the different floral visitor guilds by exploring at a broad geographical scale the reproductive spectrum of trait variation of flowering plants from plant-pollination networks. First, we investigate what are the major axes of reproductive trait variation and trait correlations for the different plant species. Second, we investigate the association between the plant species' position in the multidimensional trait-space and the different floral visitor guilds with the help of qualitative (presence-absence of interaction) and quantitative (number of visits) information about plant-pollinator interactions. Finally, we investigate how the main axes of trait variation influence plant species' functional roles in the pollination network using a set of complementary interaction network metrics (i.e., number of visits, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see table S1 and figure S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation across different biological communities. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies {[}e.g., \protect\hyperlink{ref-carvalheiro2014}{42},\protect\hyperlink{ref-olesen2007}{43},\protect\hyperlink{ref-fortuna2010}{44}{]} or available in online archives {[}e.g., Mangal \protect\hyperlink{ref-poisot2016}{45},~The Web of Life \protect\hyperlink{ref-fortuna2010}{44}{]}. In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

We selected a total of 19 different quantitative and categorical functional traits that comprised reproductive traits and plant form and size related traits (see table 1). Reproductive traits consisted of floral traits and reproductive biology traits: (i) floral traits included traits related to the size of floral organs (e.g., style length), floral display (e.g., flower number) or floral rewards (e.g., pollen quantity) while (ii) reproductive biology traits indicated the reproductive system of the plant (i.e., breeding, mating and compatibility system). The remaining traits were plant size and form related traits that are commonly used to characterize the fast-slow continuum of plant trait variation (i.e., plant height, lifespan and life form). For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 possible cells considering all traits and plant species (20 columns × 1,506 species), we were able to fill 24,341 cells (80.8\% of the dataset, see figure S2 for missing values information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life {[}\protect\hyperlink{ref-smith2018}{47},\protect\hyperlink{ref-jin2019}{48}{]}.

@@

To avoid the loss of relevant ecological information, we imputed trait missing values with the help of the function \emph{missForest} {[}\protect\hyperlink{ref-stekhoven2012}{49}{]} which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} {[}\protect\hyperlink{ref-penone2014}{50}{]}. To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} {[}\protect\hyperlink{ref-santos2018}{51}{]} based on a previous conceptual framework that considers phylogenetic eigenvectors {[}\protect\hyperlink{ref-diniz-filho2012}{52}{]}. We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (figure S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without imputed values and we found consistent similar results on both analyses (Fig S3 and Fig S4).

@@

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers {[}\protect\hyperlink{ref-serneels2008}{53},\protect\hyperlink{ref-legendre2012}{54}{]}, we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as recommended for principal component analysis {[}\protect\hyperlink{ref-legendre2012}{54}{]}. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation {[}\protect\hyperlink{ref-abdi2010}{56}{]}. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

\caption{\textbf{Figure 2.} Location of the different qualitative traits on the trait space. The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f). The solid arrows indicate the direction of variation of the different quantitative traits showed in figure 1: flower number (FN), plant height (PH), style length (SL), flower size (FS), ovule number (ON) and the level of autonomous selfing (S).}\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  a4paper,
]{article}
\title{\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shapes interactions with floral visitors}
\author{}
\date{\vspace{-2.5em}}

\usepackage{amsmath,amssymb}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={ Covariation among reproductive traits in flowering plants shapes interactions with floral visitors},
  colorlinks=true,
  linkcolor={RoyalBlue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={RoyalBlue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\pagestyle{plain}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno} % add 
\linenumbers % turns line numbering on

@@

Globally, plants display enormous variation in life history strategies and trait combinations. However, evidence suggests that evolutionary and physiological constraints limit the number of plant ecological strategies. Although there have been recent advances in understanding correlations among plant traits, reproductive traits are rarely considered, despite their key role in shaping plant life-history strategies and interactions with floral visitors. Here, using a global dataset of 16 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with floral visitors. We show that over 50\% of all trait variation is explained by the first two reproductive axes, which represent the negative correlation between flower number and flower size, and the positive correlation between pollinator dependence and floral display size. In empirical plant-pollinator networks, the plant position within these axes was associated with the number of interacting partners and the visitation rates of the distinct floral visitor guilds, but the unexplained variability found highlights the need to incorporate other factors along with reproductive traits to fully explain plant-pollinator interactions. Our study identifies the major reproductive trait correlations in flowering plants and their role in shaping networks of plant-pollinator interactions at a global scale.

@@

Flowering plants display an astonishing diversity of reproductive structures {[}\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}{]} that shape their interactions with animal pollinators {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}{]}. The variety of reproductive strategies used by different plant species, from specialized reliance on a single pollinator, to full self-pollination, is incredibly rich. However, not all reproductive trait combinations in plants are possible due to evolutionary and ecological constraints {[}\protect\hyperlink{ref-stearns1989}{5},\protect\hyperlink{ref-agrawal2020}{6}{]}. Despite recent advances in theoretical and empirical understanding of correlations between plant reproductive traits at large ecological scales {[}\protect\hyperlink{ref-salguero2016}{7}--\protect\hyperlink{ref-song2022}{11}{]}, studies typically focus on just a few traits, often in isolation. Thus, to develop a comprehensive understanding of the plant reproductive spectrum of trait variation, as done recently for root {[}\protect\hyperlink{ref-laughlin2021}{12}{]}, leaf {[}\protect\hyperlink{ref-wright2004}{13}{]} and wood {[}\protect\hyperlink{ref-chave2009}{14}{]} related traits, we require a multi-trait perspective with broad geographical coverage. Importantly, characterization of the plant reproductive spectrum could reveal key reproductive strategies among different species {[}\protect\hyperlink{ref-agrawal2020}{6}{]} and improve knowledge of associations between specific reproductive traits and pollinators {[}\protect\hyperlink{ref-roddy2021}{10}{]}.

With the recent availability of large trait databases {[}e.g., TRY \protect\hyperlink{ref-kattge2011}{15},COMPADRE \protect\hyperlink{ref-salguero2015}{16}{]}, there has been increased research on plant ecological strategies, which has produced identification of global patterns in plant form and function {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-diaz2016}{17}--\protect\hyperlink{ref-carmona2021}{19}{]}. However, studies that look at multiple traits tend to overlook plant reproductive traits {[}\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-evojtko2020}{20},\protect\hyperlink{ref-ruger2018}{21}{]}, and focus on trait correlations in other plant organs such as morphological and physiological leaf traits {[}\protect\hyperlink{ref-wright2004}{13},\protect\hyperlink{ref-shipley2006}{22}--\protect\hyperlink{ref-osnas2013}{24}{]}. Despite the lack of a holistic understanding that depicts reproductive trait covariation patterns, there are widely recognized associations between plant reproductive trait pairs (e.g., the negative correlation between flower size and flower number {[}\protect\hyperlink{ref-sargent2007}{25},\protect\hyperlink{ref-kettle2011}{26}{]}; the positive association between flower size and outcrossing rate {[}\protect\hyperlink{ref-goodwillie2010}{27}{]}; or the positive correlation between outcrossing rate and lifespan {[}\protect\hyperlink{ref-barrett2003}{28}--\protect\hyperlink{ref-moeller2017}{30}{]}). Although identification of these trait correlations has allowed progress toward a conceptual framework that integrates different floral trait relationships {[}\protect\hyperlink{ref-roddy2021}{10}{]}, we still lack empirical understanding of how these associations among floral and other reproductive traits (e.g., reproductive biology) shape plant life history strategies.

Trait-based approaches {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{31}{]} and trait-matching analyses {[}\protect\hyperlink{ref-stang2009}{32},\protect\hyperlink{ref-bartomeus2016}{33}{]} are powerful for exploring the underlying mechanisms driving plant-pollinator interactions. For example, plant traits can define species' network roles {[}e.g., specialists vs generalists \protect\hyperlink{ref-tur2013}{34}{]} and plant species that occupy reproductive trait space extremes are likely to be more specialized and display tighter trait-matching with pollinators {[}\protect\hyperlink{ref-junker2013}{35},\protect\hyperlink{ref-coux2016}{36}{]}. Indeed, the degree of morphological matching between plants and floral visitors often determines whether or not interactions occur, and consequently influences the broader network structure {[}\protect\hyperlink{ref-stang2009}{32},\protect\hyperlink{ref-ibanez2012}{37}{]}. However, it is still unclear how specific reproductive traits (e.g., mating or compatibility system) influence plant-pollinator associations {[}\protect\hyperlink{ref-tur2013}{34},\protect\hyperlink{ref-sargent2008}{38},\protect\hyperlink{ref-devaux2014}{39}{]}. Indeed, both morphology and mating system can determine a species' functional role within pollination networks, and the combination of several traits increases power for predicting network structure {[}\protect\hyperlink{ref-eklof2013}{40}{]}. Because the species position within the multidimensional trait space can result in different interaction patterns or adaptations to pollinators {[}see \protect\hyperlink{ref-dehling2016}{41} for a purely morphological approach{]}, the use of the reproductive trait space to explore plant-pollinator interactions could help to identify species functional roles and associations of traits with specific pollinators (i.e., pollination syndromes).

Here, we explore the spectrum of reproductive traits in flowering plants, within a global set of plant-pollinator networks, to identify reproductive trait covariation and how this influences the network of interactions between plants and floral visitors. First, we identify the major axes of reproductive trait variation and correlations among traits for plant species. Then, we investigate the associations between plant species' positions in reproductive trait space and the strength of their interactions with different floral visitor guilds within networks. Finally, we determine how the main axes of reproductive trait variation influence plant species' functional roles within pollination networks, using a range of complementary interaction network metrics.

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see table S1 and figure S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation across different ecological communities. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to benefit, to some degree, from animal pollination. Many of these networks are freely available either as published studies {[}e.g., \protect\hyperlink{ref-carvalheiro2014}{42},\protect\hyperlink{ref-olesen2007}{43},\protect\hyperlink{ref-fortuna2010}{44}{]} or in online archives {[}e.g., Mangal \protect\hyperlink{ref-poisot2016}{45},The Web of Life \protect\hyperlink{ref-fortuna2010}{44}{]}. In total, our network dataset constituted 60 weighted (number of visits) and four unweighted (interaction presence-absence) networks, each sampled at a unique location and year, as well as eight meta-networks where interactions were pooled across several locations and multiple years.

@@

We selected a total of 19 quantitative and categorical functional traits that included reproductive traits and traits related with plant form and size (see table 1). Reproductive traits consisted of those related to floral and reproductive biology: (i) floral traits included those related to the size of floral organs (e.g., style length), floral display size (e.g., flower number) and floral rewards (e.g., pollen quantity), while (ii) reproductive biology traits represented the plant reproductive system (i.e., the breeding, mating and compatibility system). The remaining traits were related to plant size and form (i.e., plant height, lifespan and life form), which are commonly used to characterize the fast-slow continuum of plant trait variation {[}\protect\hyperlink{ref-salguero2016}{7}{]}. For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a matrix comprising a total of 28,614 possible cells (considering all traits and plant species: 19 columns × 1,506 species), we were able to fill 23,354 cells (81.6\% of the dataset, see figure S2 for missing value information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

We calculated the phylogenetic distance between all plant species considered using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life {[}\protect\hyperlink{ref-smith2018}{47},\protect\hyperlink{ref-jin2019}{48}{]}.

@@

To avoid excluding species with trait missing values, we imputed trait missing values with the help of the function \emph{missForest} {[}\protect\hyperlink{ref-stekhoven2012}{49}{]} which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors from principal component analysis of the phylogenetic distance (PCoA), which has been shown to improve the performance of \emph{missForest} {[}\protect\hyperlink{ref-penone2014}{50}{]}. To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} {[}\protect\hyperlink{ref-santos2018}{51}{]} based on a previous conceptual framework that considers phylogenetic eigenvectors {[}\protect\hyperlink{ref-diniz-filho2012}{52}{]}. We conducted two different imputations, one for the full set of species (1,506 species: 5.79\% of missing values), and excluded nectar and pollen traits because of the high percentage of missing values (figure S2), and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species: 8.01\% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without imputed values and results for the two datasets were consitent (Fig S3 and Fig S4).

@@

We explored the association between the different quantitative plant traits using a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width, respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Specifically, due to the high sensitivity of dimensionality reduction to outliers {[}\protect\hyperlink{ref-serneels2008}{53},\protect\hyperlink{ref-legendre2012}{54}{]}, we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed all continuous variables to reduce the influence of extreme values and then z-transformed (X= 0, SD=1) these variables so that they were within the same numerical range, as recommended for principal component analysis {[}\protect\hyperlink{ref-legendre2012}{54}{]}. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation using an ANOVA and Tukey test. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode ``covariance'' because values for each variables were on the same scale following transformation {[}\protect\hyperlink{ref-abdi2010}{56}{]}. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity data. For this, we included all quantitative traits included in the main pPCA, plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} and used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

\caption{\textbf{Figure 2.} Location of the different qualitative traits in the trait space. The figure panel is composed of the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f). The solid arrows indicate the direction of variation of the different quantitative traits shown in figure 1: flower number (FN), plant height (PH), style length (SL), flower size (FS), ovule number (ON) and the level of autonomous selfing (S).}\label{fig:unnamed-chunk-3}
\end{figure}
